---
title: "Data Transformation"
author: "Sean O'Neal"
date: "9/18/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 5.1 Introduction

### 5.1.1 Prerequisites

```{r}
library(tidyverse)
library(nycflights13)
library(Lahman)
```

### 5.1.2 nycflights13

```{r}
flights
```

### 5.1.3 dplyr

verbs

filter
arrange
select
mutate
summarize

verb(database, whatToDo)

# 5.2 Filter rows with filter()

Flights that occurred in Jan 1st

```{r}
filter(flights, month == 1, day == 1 )
```

```{r}
(jan1 <- filter(flights, month ==1, day ==1))
```

```{r}
(dec25 <- filter(flights, month == 12, day ==25))
```

### 5.2.1 Comparisons 

==, >, >=, <, <=, !

```{r}
filter(flights, month == 1)
```

```{r}
cos(pi/2) ==0
```

```{r}
near(cos(pi/2), 0)
```

### 5.2.2 Logical Operators

or ---------> |
and -----> &
not ------> !

flights that departed in November **or** December

```{r}
filter(flights, month ==11 | month== 12)
```

```{r}
filter(flights, month == 5 | month ==6 | month == 7 | month == 8)
```

```{r}
filter(flights, month %in% c(5,6,7,8))
```

Flights that departed in November **and** arrived in December

```{r}
#tbd
```

```{r}
filter(flights, !(dep_delay > 120 | arr_delay >120))
```

##### 5.2.3 Missing values

One important feature of R that can make comparison tricky are missing values

```{r}
100 == NA
2 > NA
x <- NA
is.na(x)
```

[1] NA
[1] NA
[1] FALSE

How do I get rid of NA's on my dataset?

```{r}
filter(flights, is.na(dep_time))
filter(flights, !is.na(dep_time))
withoutNA <- filter(flights, !is.na(dep_time))
withoutNA

filter(withoutNA, is.na(dep_time))
```

#### 5.2.4 Exercises 

```{r}
flights
```

> 1. Find all flights that

had an arrival delay of two or more hours

```{r}
filter(flights, arr_delay >= 120)
```

Flew to Houston (IAH or HOU)

```{r}
filter(flights, dest == 'IAH' | dest == 'HOU')
filter(flights, dest %in% c('IAH', 'HOU'))
```

Were flown by the carriers United, American, and Delta
```{r}
filter(flights, carrier %in% c('UA', 'AA', 'DL'))
```

Were Flown in the months of july, august, and september

```{r}
filter(flights, month %in% c(7,8,9))
```

Arrived 2 hrs late and left on time

```{r}
filter(flights, arr_delay > 120 & dep_delay <= 0)
```

```{r}
filter(flights, dep_delay >= 60, arr_delay < 30)
```

Departed between midnight and 6 am

```{r}
filter(flights, between(dep_time, 00, 600))
```




> 2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenge?




> 3. how many flights have a missing dep_time? what other variables are missing?

```{r}
filter(flights, is.na(dep_time))
```

### 5.5 Add new variables with mutate()

```{r}
flights_narrow <- select(flights, year:day, ends_with('delay'), distance, air_time)
```

Mutate function

```{r}
mutate(flights_narrow, gain = dep_delay - arr_delay, speed = distance / air_time * 60)
```

```{r}
mutate(flights_narrow, gain = dep_delay - arr_delay, speed = distance / air_time * 60, gain_per_hour = gain / air_time)
```

Use transmute if you only want to keep the new variables

```{r}
transmute(flights, gain = dep_delay - arr_delay, speed = distance / air_time * 60, gain_per_hour = gain / air_time)
```

```{r}
select(flights, year, month, day)
select(flights, year:day)

select(flights, -(year:day))
select(flights, dep_time:time_hour)
```

Helper functions to select columns

```{r}
select(flights, starts_with('dep'))
select(flights, ends_with('delay'))
select(flights, contains('_'))
```

#5.6 Grouped summaries with summarise()

```{r}
summarize(flights, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_day <- group_by(flights, year, month, day)

summarize(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_carrier <- group_by(flights, carrier)

summarize(by_carrier, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_airport <- group_by(flights, origin)

summarise(by_airport, delay = mean(dep_delay, na.rm = TRUE))
```

#### 5.6.1 Combining multiple operations with the pipe

Group flights by destination:

```{r}
by_dest <- group_by(flights, dest)

by_dest
```

summarize the group using average distance, average delay, and count:

```{r}
delays <- summarise(by_dest, 
count = n(), 
dist = mean(distance, na.rm = TRUE),
delay = mean(arr_delay, na.rm = TRUE)
)

delays 
```

Make a plot of distance vs delay with count mapped to the size aesthetic

```{r}
ggplot(data = delays) + 
  geom_point(mapping = aes(x = dist, y = delay, size = count))
```

remove noisy points

```{r}
delays <- filter(delays, count > 20, dest != 'HNL')
delays
```

visualize again

```{r}
ggplot(data = delays) +
  geom_point(mapping = aes(x = dist, y = delay, size = count))
```

clean up the plot

```{r}
ggplot(data = delays, mapping = aes(x = dist, y = delay)) +
  geom_point(mapping = aes(size = count), alpha = .3) +
  geom_smooth(se = FALSE) #se = False - delates the standard error band
```

```{r pipe}
delays <- flights %>%  
  group_by(dest) %>%
  summarize(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>%
  filter(count > 20, dest != 'HNL')

ggplot(data = delays, mapping = aes(x = dist, y = delay)) +
  geom_point(mapping = aes(size = count), alpha = .3) +
  geom_smooth(se = FALSE)

```

##### 5.6.2 Missing Values

you may have wondered about the na.rm argument, what happens if we dont set it?

```{r}
flights %>%
  group_by(year, month, day) %>%
  summarize(delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
not_cancelled <- flights %>%
  filter(!is.na(arr_delay) & !is.na(dep_delay))
```

#### 5.6.3 Counts

```{r}
delays <- not_cancelled %>%
  group_by(tailnum) %>%
  summarize(delay = mean(arr_delay))

delays

arrange(delays, desc(delay))
```

Wow, there are planes with a huge amount of delay time

```{r}
delays <- not_cancelled %>%
  group_by(tailnum) %>%
  summarize(
    delay = mean(arr_delay),
    count = n()
  )

arrange(delays, desc(delay))

ggplot(data = delays) +
  geom_point(mapping = aes(x = count, y = delay), alpha = 1/3)
```

**whenever you plot a mean or other summary vs group size, you'll see that the variation decreases as the sample size increases**

```{r}
delays %>%
  filter(count > 25) %>%
  ggplot() + 
  geom_point(mapping = aes(x = count, y = delay), alpha = .1)
```

```{r}
Batting
```

```{r}
batting <- Batting

batters <- batting %>%
  group_by(playerID) %>%
  summarize(
    ba = sum(H, na.rm = TRUE) / sum(AB, na.rm = TRUE),
    ab = sum(AB) #at bats
  )

batters %>%
  filter(ab > 100) %>%
  ggplot(mapping = aes(x = ab, y = ba)) + 
  geom_point() +
  geom_smooth(se = FALSE)

```







