---
title: "Data Visualization"
author: "Sean O'Neal"
date: "8/21/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Visualization

### 3.1 Introduction

```{r}
# we must load the tidyverse library every session

library(tidyverse)
```

### 3.2 First Steps 

QUESTION: Do cars with big engines use more gas than cars with small engines?

##### 3.2.1 the mpg data frame

A data frame is a rectangular collection of variables (in the columns) and observations (in the rows). mpg contains observations collected by the US Environmental Protection Agency on 38 models of car.

```{r}
mpg

```
##### 3.2.2 Creating a ggplot

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))

```
##### Exercises 

>1. Run ggplot(data = mpg). What do you see?

```{r}
ggplot(data = mpg)
```
I see a large blank grey slate

>2.How many rows are in mpg? How many columns?

There are 234 rows and 11 columns

>3. What does the drv variable describe? Read the help for ?mpg to find out.

The drv variable describes the drive-train of the car, whether it's a front-wheel drive, rear-wheel drive, or all-wheel drive car.

>4. Make a scatterplot of hwy vs cyl.

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = hwy, y = cyl))
```

>5. What happens if you make a scatterplot of class vs drv? Why is the plot not useful?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = class, y = drv))
```
This plot is not useful because the class of car is not always dependent on the drive train and the variables are categorical.

### Aesthetic mappings

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, size = class))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

```{r}
ggplot(data = mpg) +geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

##### 3.3.1 Exercises 

> 1. What’s gone wrong with this code? Why are the points not blue?

This code is incorrect because the "color = "blue"" is supposed to be inside of the second group of parentheses.
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

> 2. Which variables in mpg are categorical? Which variables are continuous?

The categorical variables are manufacturer, model, trans, drv,fl, and class. The continuous variables are displ, year, cyl, cty, and hwy.
```{r}
?mpg
```

> 3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

For colour, continuous variables change shades of the same colour while categorical changes the entire colour. For size, the variables react in the same way but there is more data for the continuous variables than the categorical. For shape, continuous variables cannot be be mapped to shape while categorical variables can.
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, shape = drv))
```

> 4. What happens if you map the same variable to multiple aesthetics?

When you map the same variable to multiple aesthetics the variable that is already mapped is mapped again but as an aesthetic. 
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = hwy))
```

> 6. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?

When you map something other than a variable name, the graph takes the data and makes it conform to the specification that you typed. For this example, all displacements larger than 5 litres are coloured red and all displacements smaller than 5 litres are coloured blue.
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
```

### 3.5 Facets

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_wrap(~ class, nrow = 2)
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(drv ~ cyl)
```

> 1. What happens if you facet on a continuous variable?

It turns the continuous variable into categories for the graphs. 
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_wrap(~ cyl, nrow = 2)
```


> 2. What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?

You see that 5 cylinder 4 wheel drive and rear wheel drive cars, as well as 4 cylinder rear wheel drive cars are very rare and/or nonexistent. 
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(drv ~ cyl)
```


> 3. What plots does the following code make? What does . do?

The "." on the left side organized the data into columns while the "." on the right side organizes the data into rows.
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(drv ~ .)
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(. ~ cyl)
```


> 4. What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

The advantages of using faceting are that it organizes the variables into smaller separate graphs based on the criteria you define. The balance would change with a larger data set because it is easier to discern the data in facets as opposed to multiple colours that may look similar. 

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_wrap(~ class, nrow = 2)
```


> 5. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?

nrow and ncol changes the number of rows and columns. The other options are scales, shrink, labeller, as.table, switch, drop, dir, and strip.position. Facet_grid doesnt have nrow and ncol arguments because they don't apply to that type of command. 

```{r}
?facet_wrap
```

> 6. When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?

You should put the variable with more unique levels in the columns because the facet_grid command has an easier time organizing and graphing larger variables in columns

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(cyl ~ cty)
```

### 3.6 Geometric objects

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
```

```{r}
ggplot(data = mpg) +geom_smooth(mapping = aes(x = displ, y = hwy, color = drv))
```

```{r}
ggplot(data = mpg) +geom_point(mapping = aes(x = displ, y = hwy)) + geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = drv)) + geom_smooth(mapping = aes(x = displ, y = hwy, color = drv))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +geom_point() + geom_smooth()
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + geom_smooth() + geom_point(mapping = aes(color = class))
```

## Excercises 1-5 

> 1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

To draw a line chart, you would use geom_abline. To make a box plot, you would use geom_boxplot, to make a histogram, you would use geom_histogram, and to make an area chart you would use geom_area.

> 2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

I predict that this will look like a normal chart with both the line and the points coloured by drv, but the se takes out the standard error. This was correct.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + geom_point() + geom_smooth(se = FALSE)
```


> 3. What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?

Show.legend = false drops the legend from the graph. When you remove it, the legend appears again. You used this earlier in the chapter in order for the graphs to not get too complicated. 

```{r}
ggplot(data = mpg) +geom_point(mapping = aes(x = displ, y = hwy, show.legend = FALSE))
```


> 4. What does the se argument to geom_smooth() do?

The se argument computes the standard error of the mean. 

> 5. Will these two graphs look different? Why/why not?

No they will not because it is the same exact data and commands but they are spread out over the different chunks of code. 

### 3.7 Statistical transformations

```{r}
diamonds
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut))
```

### 3.8 Position adjustments

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, color = cut))
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = cut))
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = clarity))
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + geom_bar(alpha = 0.2, position = "identity")
```

```{r}
ggplot( data = diamonds, mapping = aes(x = cut, color = clarity)) + geom_bar(fill = NA, position = "identity")
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + geom_bar(position = "fill")
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + geom_bar(position = "dodge")
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")
```

##### 3.8.1 Exercises

Notebook check - exercises and notes 

Project - start next week / due the week after

> 1. The problem with this plot is that the points will be automatically calculated into averages, so you could improve it by adding the "jitter" adjustment to add some randomness to the graph in order for it to be more realistic.


> 2. The parameters that controls the amount of jittering are width and height

```{r}
?geom_jitter
```


> 3. Geom_jitter makes the points move around a little more so that you can see all of the various points if they are overlapping. Geom_count counts the number of overlapping points in one area and maps the count to point area. 

```{r}
?geom_count
```


> 4.

```{r}
ggplot(data = mpg, mapping = aes(x = displ)) + geom_boxplot()
```










